---
title: "lab5"
author: "Andrew Do"
date: "June 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Help for Packages

You can see what an entire package has to offer with
```{r, eval = F}
help(package = "stringr")
```


## Counting number of vowels

Consider the following vector `letrs` which contains various letters:
```{r}
# Garbage string
letrs <- "gfngenodgsyctyviqyyigeinyngbkwjdjqkrrollhpxdkfldlytympmcgmjynihusebtq"
```

If you were to count the number of vowels in `letrs` you would get the following counts:

- a: 0
- e: 3
- i: 4
- o: 2
- u: 1

This exercise will guide you through how to arrive at this frequency table.

```{r}
# Create a character vector of vowels.  Give it an appropriate name.
vowels <- c("a", "e", "i", "o", "u")

# Use str_count to get the counts.  Assign this an appropriate name.
vcount <- str_count(letrs, vowels)

# Give each element of the vector a name using the names function.
# The result should look like:
# a e i o u
# 0 3 4 2 1

names(vcount) <- vowels
vcount
```

## Exploring new functions: str_pad
Write a short description of what each of the following calls of `str_pad` is doing.  The last three examples were found at the bottom of the function's documentation.

1. In this call of `str_pad`, the string "x" is tacked onto the left side of the string "a" until the string has 4 characters.
```{r}
str_pad("a", width = 4, side = "left", pad = "x")
```

2. In this call of `str_pad`, the space string " " is tacked onto the left side of the string "a" until the string has 3 characters.
```{r}
str_pad("a", width = 3, side = "left")
```

3. In this call of `str_pad`, spaces are added to the three elements of the vector until they reach 10 characters.  The 10 has been recycled to match the length 3 character vector.
```{r}
str_pad(c("a", "abc", "abcdef"), 10)
```

4. In this call of `str_pad`, the default padding character, the space, has been added to the left side of the string "a" until it reached 5, 10, and 20 characters.  The "a" was recycled to match the length of `c(5, 10, 20)`.
```{r}
str_pad("a", c(5, 10, 20))
```

5. In this call of `str_pad`, the string "a" was padded with hyphens, underscores, and " " to produce a length 10 string.  The "a" and the 10 were recycled until they reached length 3.
```{r}
str_pad("a", 10, pad = c("-", "_", " "))
```

## Regular Expressions

We read in the data from the csv file and extract the names column
```{r}
flags <- read_csv("flags.csv")
countries <- flags$name
```

You may have noticed that in the `countries` vector, words in a country's name are separated by `-`.  Use `str_replace_all` to replace the hyphens with spaces.  For example `"Vatican-City"` should be replaced with `"Vatican City"`

```{r}
countries <- str_replace_all(countries, "-", " ")
```

Use `str_detect` to see if there are any countries with the world "island" in them, ignoring case.  You may also want to look up how to use `regex` as a way to ignore case.

```{r}
any(str_detect(countries, regex("island", ignore_case = T)))
```

Use `str_subset` to find all countries that begin with a capital `A`.
```{r}
str_subset(countries, "^A")
```

Find all countries that end with `stan`
```{r}
str_subset(countries, "stan$")
```

Use `str_replace_all` to replace the `$`'s in `baws` with `s`'s.  Recall that the `$` is a metacharacter.
```{r}
baws <- "Bo$$"
str_replace_all(baws, "\\$", "s")
```

Use a character class (denoted by square brackets) to find all countries beginning with a vowel and ending with a consonant.
```{r}
str_subset(countries, "^[AEIOU].*[^aeiouAEIOU]$")
```




