---
title: "lab3"
author: "Andrew Do"
date: "June 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing the Data
Run install the package `readr` by running `install.packages("readr")` in the console.

The code below reads in the `tents.csv` comma-separated file.  The `col_types` argument is telling `R` what the data type is in each column.  See `?read_csv` if you're curious.
```{r}
library(readr)
tents <- read_csv("tents.csv", col_types = "ccnniccc")
```

### Questions
Answer the following questions programatically.  Some questions ask for you to explain what you see---you may answer this as text outside a code block.

1. Run `str()` on `tents`.  What information does it tell you? This is a useful function to know.
This function gives the structure of an R object --- names it contains, what class it is, data types it holds, etc.
```{r}
str(tents)
```

2. What class is `tents`?
```{r}
class(tents)
```

3. How many rows are in `tents`?
```{r}
nrow(tents)
```

4. How many columns are in `tents`?
```{r}
ncol(tents)
```

5. What are the names of the columns?
```{r}
names(tents)
```

6. Display the first 5 rows.
```{r}
head(tents, 5)
```

7. Display the last 4 rows.
```{r}
tail(tents, 4)
```

8. Find out how to use `quantile` to find the interquartile range (75th percentile - 25th percentile) of the weights of the tents.
```{r}
iqr <- quantile(tents$weight, .75) - quantile(tents$weight, .25)
names(iqr) <- NULL
iqr

# You can also use unname to remove the name and diff to find the difference
# unname(diff(quantile(tents$weight, c(.25, .75))))
```

9. Find the standard deviation of the heights of the `tents`.  Does `R` use `n` or `n-1` in the denominator when calculating standard deviation?
By default, `R` uses the unbiased estimate of standard deviation, so the denominator is `n-1`.
```{r}
sd(tents$height)
```

10. The weight is given in grams.  Add a new column to the data frame called `weight_lbs` that contains the weight of the tents in pounds.  Remember to overwrite `tents` with your update.
```{r}
tents$weight_lbs <- tents$weight * 0.00220462
```

11. The height is given in centimeters.  Add a new variable `height_in` to the data frame that contains the height of the tents in inches. Remember to overwrite `tents` with your update.
```{r}
tents$height_in <- tents$height * 0.393701
```

12. How many tents cost more than $400?
```{r}
sum(tents$price > 400)
```

13. Subset the data frame so that it only contains tents that are valued $300 or less.  Assign the subsetted data frame the name `so_cheap_tents`.
```{r}
so_cheap_tents <- tents[tents$price <= 300, ]
```

14. Of the tents that are more than $300, what percent are `big-agnes` tents?
```{r}
with(tents, 100 * mean(brand[price > 300] == 'big-agnes'))
```

15. Select the `bestuse` column of `tents` in four different ways: using `$`, `[[]]`, position number and `[]`, and variable name and `[]`. 
```{r}
# These two return vectors
tents$bestuse
tents[["bestuse"]]

# These two return single-column data frames
tents[6]
tents["bestuse"]
```

16. What is the name of the tent with the best height-to-weight ratio?
```{r}
with(tents, name[which.max(height / weight)])
```

17. Run `table` with two arguments, `seasons` and `capacity` (this should be one table).  What did it do?
This call tabulates a contingency table between `seasons` and `capacity`, displaying counts with each value-pair.
```{r}
with(tents, table(seasons, capacity))
```

18. What type of tent (bestuse) is the most common? Your answer should return only the use and not the count.
```{r}
names(sort(table(tents$bestuse), decreasing = T))[1]
```

19. Create a data frame called `kelty` with only data on kelty tents.  Do the same for `big-agnes`.  Then, combine the two data frames.
```{r}
# You shouldn't actually ever do something like this.  This is purely an
# exercise.
kelty <- tents[tents$brand == "kelty", ]
big_agnes <- tents[tents$brand == "big-agnes", ]
rbind(kelty, big_agnes)
```

